---
title: "STAT_450_Project"
author: "Noah Gallagher"
date: "4/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Let us first read in the data and take a look at the variables.

```{r}
# Reading in data
vegas <- read.csv('vegas2.csv',sep = ',')

# Looking at variables
str(vegas)
```

## Generate Summary Statistics

Here we can dive deeper into the variables and generate some summary statistics.

```{r}
library(skimr)
skim(vegas)

#create subset of numeric elements
numeric_subset<- vegas2%>%
  select(Nr..reviews, Nr..hotel.reviews, Helpful.votes, Score, Hotel.stars, Nr..rooms, Member.years)

# get col means
colMeans(numeric_subset)
```

# Generating Plots

```{r}
library(gridExtra)
library(tidyverse)
# Plotting Hotel Reviews
hotel_plot = ggplot(vegas, aes(x=Nr..hotel.reviews)) +
            geom_histogram(bins=15) +
            labs(x='Reviews', y='count')

#hotel_plot

# Plotting scores
scores_plot = ggplot(vegas, aes(x=Score)) +
            geom_histogram(bins=5) +
            labs(x='Scores', y='count')
#scores_plot

# Plotting Number of rooms
rooms_plot = ggplot(vegas, aes(x=Nr..rooms)) +
            geom_histogram(bins=5) +
            labs(x='Number of Rooms', y='count')
#rooms_plot

# Merging
grid.arrange(hotel_plot, scores_plot, rooms_plot, ncol=2, nrow=2)

# Plotting Pool Presence
pool_plot = ggplot(vegas, aes(x=Pool)) +
            geom_bar() +
            labs(x='Presence of Pool', y='count')
#pool_plot

# Plotting Gym Presence
gym_plot = ggplot(vegas, aes(x=Gym)) +
            geom_bar() +
            labs(x='Presence of Gym', y='count')
#gym_plot

# Plotting Spa Presence
spa_plot = ggplot(vegas, aes(x=Spa)) +
            geom_bar() +
            labs(x='Presence of Spa', y='count')
#spa_plot

# Plotting Tennis Court Presence
tennis_plot = ggplot(vegas, aes(x=Tennis.court)) +
            geom_bar() +
            labs(x='Presence of Tennis Court', y='count')
#tennis_plot

# Plotting Casino Presence
casino_plot = ggplot(vegas, aes(x=Casino)) +
            geom_bar() +
            labs(x='Presence of Casino', y='count')
#casino_plot

# Plotting Free Wifi
wifi_plot = ggplot(vegas, aes(x=Free.internet)) +
            geom_bar() +
            labs(x='Presence of Wifi', y='count')
#wifi_plot

# Merging
grid.arrange(pool_plot, gym_plot, spa_plot, tennis_plot, casino_plot, wifi_plot, ncol=3, nrow=2)
```




Questions:
  1) Can we determine if there is a significant difference between the stars a hotel has and their reviews?
  
  Testing: Ho: Mu-Score = Mu-Stars vs. Ha: Mu-Score != Mu-Stars
```{r}
table(vegas$Hotel.name)
attach(vegas)
Y = cbind(Score, Hotel.stars)
grp = factor(Hotel.name)
fit = manova(Y ~ grp)
summary(fit, test = 'Wilks')
```

ANOVA test - analysis of individual variables
```{r}
# AOV test for Score
summary(aov(Score ~ grp))

# AOV test for Hotel.stars
summary(aov(Hotel.stars ~ grp))

# F stat
g=21
p=2
N=504
alpha=0.05
qf(1-(alpha)/p, N-g, g-1)
#F = 2.10

```
For Hotel Score: We Reject Ho meaning, all hotel ratings differ significantly among the Hotels and there is no correlation between them.

For Hotel Stars: We Reject Ho, meaning all hotel Stars differ significantly among the Hotels.



  2) Can we determine if there is a significant difference between tourist reviews and USA resident reviews for hotels?
  
  Testing: Ho: Mu-Score_tourist = Mu-Score_resident vs. Ha: Mu-Score_tourist != Mu-Score_resident
```{r}
# Lets use ANOVA test 
attach(vegas)
summary(aov())


```

  Sub Question) Can we assess the significance of individual amenities
```{r}
# AOV Test for each Amenity
vegas2<-na.omit(vegas)
attach(vegas2)
grp = factor(Hotel.name)


# Pool
summary(aov(Pool ~ grp),success='YES')
```
  Sub Question) Is there a difference in Reviews based on what amenities are offered?


  3) What is the most popular country that tourists come from?
```{r}
country_sums <- vegas%>%
  group_by(User.country)%>%
  summarize(count=n())%>%
  arrange(desc(count))
country_sums
```
We have determined that the most popular country that tourists come from is USA, followed by UK.

  4) What is the highest rated hotel on the Las Vegas Strip?
```{r}
highest_rated <- vegas%>%
  group_by(Hotel.name)%>%
  summarise(
    num_reviews = sum(Nr..hotel.reviews),
    total_score = sum(Score),
    avg_rating = mean(Score)
  )%>%
  arrange(desc(avg_rating))
highest_rated
```
Thus, we can conclude that the Wynn Las Vegas is the highest rated hotel on the Las Vegas Strip.


  5) What is the most popular traveler type?
```{r}
popular_traveler <- vegas%>%
  group_by(Traveler.type)%>%
  summarize(count = n())%>%
  arrange(desc(count))

popular_traveler
```
The most popular traveler type that visit Las Vegas are Couples, followed by families.

## Maybe we can visualize this
  6) What amenities are normally included for all hotels? 
  
  We know that there are 24 records for each Hotel (i.e. 24 reviews by different customers for each hotel). Let 1 = yes, 0 = no.
  Steps:
    -We first create a dataset that contains boolean values for each amenity that a hotel offers to its guests.
    -We then find the number of total hotels in the dataset.
    -We then find the proportions of amenities that the hotel offers.
```{r}
# Lets see what hotels have certain amenities
normal_amenities <- vegas%>%
  group_by(Hotel.name)%>%
  summarize( 
    pools = sum(Pool == 'YES')/24, #we divide by 24 since there are 24 entries for each hotel
    gyms = sum(Gym == 'YES')/24,
    tennis = sum(Tennis.court == 'YES')/24,
    spas = sum(Spa == 'YES')/24,
    casinos = sum(Casino == 'YES')/24,
    wifi = sum(Free.internet == 'YES')/24
  )
normal_amenities

# Get the number of hotels in dataset (21 hotels)
count_hotels <- vegas%>%
  summarize(num_hotel = length(unique(Hotel.name)))
count_hotels

# Lets get the proportions of amenities
hotel_props<- normal_amenities%>%
  summarize(
    prop_pools = sum(pools) / 21,
    prop_gym = sum(gyms) / 21,
    prop_tennis = sum(tennis) / 21,
    prop_spas = sum(spas) / 21,
    prop_casinos = sum(casinos) / 21,
    prop_wifi = sum(wifi) / 21
  )

hotel_props
```
Thus we can conclude:
  - 95.24% of hotels have Pools
  - 95.24% of hotels have Gyms
  - 23.81% of hotels have Tennis Courts
  - 76.19% of hotels have Spas
  - 90.48% of hotels have Casinos
  - 95.24% of hotels have Free Internet
  
## Correlation Between Hotel Reviews and Stars

```{r}
vegas2<- vegas%>%
  mutate(
    Hotel.stars = as.numeric(Hotel.stars),
    Score = as.numeric(Score)
  )
# Removing NA's
vegas2<-na.omit(vegas)

attach(vegas2)

# Correlation
cor(Score, Hotel.stars)

# Fitting a line to Score and Stars
ggplot(vegas2, aes(x = Score, y = Hotel.stars))+
  geom_point(shape = 1) +
  geom_smooth(method = lm) +
  geom_jitter(shape = 1)

```
  The Correlation is a weak positive correlation.
  
## Scatter Plot Matrix for Score and Stars
```{r}
subset <- vegas2%>%
  select(Score, Hotel.stars)

pairs(subset, lower.panel = NULL)


# check proportion of points within the prediction ellipse
S = cov(subset)
xbar = colMeans(subset)
D2 = mahalanobis(subset, xbar, S)
p = ncol(subset)
#qchisq(.95, p)

sum(D2 <= qchisq(.95, p))/nrow(subset)
```
The proportion of points within the prediction ellipse is 96.23%


## Correlation between all numeric variables

```{r}
numeric_subset<- vegas2%>%
  select(Nr..reviews, Nr..hotel.reviews, Helpful.votes, Score, Hotel.stars, Nr..rooms, Member.years)

cor(numeric_subset)
pairs(numeric_subset)
cov(numeric_subset)
colMeans(numeric_subset)

```

## Most Frequent Period of Stay

```{r}
period_stay <-vegas%>%
  select(Period.of.stay)%>%
  group_by(Period.of.stay)%>%
  summarize(
    num_guests = n()
  )%>%
  arrange(desc(num_guests))
period_stay
```
We can conclude that most guests decide to vists Vegas between the months of March to May.

## Is there a correlation between the Period of Stay and the Traveler Type?

```{r}




#Couples	214			
#Families + Friends = 110	+ 82 = 	192
#Business	74			
#Solo	24	
```